There are various tools and models in the area of probabilistic verification, especially in the analysis of non-Markovian Model. One of those model is stochastic Petri Nets. Oris is a tool for the analysis of stochastic Petri nets \cite{Bucci2010}. Oris allows to analyze Petri net with plenty of time distributions for each event such as exponential, Dirac, Uniform or instant. The expressiveness is sufficient to work on Petri nets that are equivalent to an ACTMC. This tool compute the Steady states probability as well as well as the transient state probability which is state probability at a given time. 

A less expressive model is the CTMC defined in Subsection \ref{CTMC_def}, PRISM\cite{KNP11} and Storm \cite{DBLP:journals/corr/DehnertJK017} are both well known tools to analyze CTMC. The engine of those tools are quite efficient for analysis due to the memoryless property of the exponential distribution. Phase-type (PH) approximation can be used to approach a non-Markovian model. The downside of PH approximation is the increase of the number of states which can slow down the process of analysis.  PH Fitting can be challenging depending on the PH distribution\cite{DBLP:journals/corr/KorenciakKR14}.
There are kinds of distributions which require a lot of phases to be accurately approached by PH fitting. One of those distributions is a shifted exponential distribution, which can be seen as a sequence of a deterministic timeout and a exponential distribution. The motivation of using those distributions is to depict better some events like transmissions that can't happen immediately due to physical limits.

There are automatic tools that do the Phase-Type Fitting like HyperStar \cite{10.1007/978-3-642-40725-3_13}. The power of this tools is to work on real dataset. According the system you want to model it is indeed practical to use real data such that the model depict more precisely the system. 


An alternative of PH-approximation is to work directly with a d-CTMC which is a CTMC plus deterministic distribution. This paper \cite{DBLP:journals/corr/BaierDKKR17} depict an efficient approach to analyze ACTMC with distributions where phase-type increase significantly the number of states with reasonable precision. Those distributions are amongst shifted exponential distribution and fixed-delay distribution.

A different objective on the analysis of non-Markovian model is to synthesize optimal parameter. We can add a reward system on the model and computed the optimal parameter, like a timeout on some transitions such that the mean-payoff is $\epsilon$-optimal \cite{DBLP:journals/corr/BaierDKKR17}. The mean-payoff can be seen as the average long term efficiency of the system.